\mytitle

### 2020-04-22 paper.tex
literature remark.
literature.
Revised abstract

\title{\mytitle}
%\subtitle{Extended Abstract}
%\author{}
%\institute{}
%\author{\myauthor}
%\institute{Department of Mechanical and Industrial Engineering, University of Toronto\\
%\email{sariks@mie.utoronto.ca, ecohen@mie.utoronto.ca, jcb@mie.utoronto.ca}}

@@

\begin{abstract} 
%Wind farms play an increasingly important
%role as a source of renewable energy and 
The wind farm layout optimization (WFLO) problem requires the
selection of wind turbine positions to maximize energy production. While a number of heuristic and exact techniques have been proposed in the literature to solve WFLO, it is common to model the interaction between turbines as a quadratic objective function as, indeed, has been done in some existing optimization models. Recent advances in optimization hardware and software have both targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems. 
%It is a key problem 
%when designing a new wind farm, as it has a direct influence
%on the amount of the total produced energy. 
%In recent years there have been advances in both software technologies (e.g., the extension of commercial mixed integer linear programming solvers to handle quadratic problems) and nascent hardware development (e.g., quantum computing chips and specialized classical chips) that have relevance for the WFLO problem. In both cases, the advances have 
%Recent years saw 
%advances in software and hardware optimization technologies, such as Gurobi's quadratic solver 
%and Fujitsu's digital annealer (DA), respectively. 
%These technologies create an opportunity to improve 
%over current wind farm optimization solutions
%both in terms of solution quality and solve time. 
%In order to exploit these technologies, 
%we must represent the WFLO problem as 
%a quadratic optimization model. Specifically, in order 
%to apply a software-based optimization
%solution (Gurobi), 
In this paper, we 
introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
%Subsequently, we provide the unconstrained counterpart of the QCOP, 
%namely the quadratic unconstrained binary optimization (QUBO), which  
%is the underlying representation for quantum-inspired 
%optimization hardware such as the DA. 
A thorough empirical 
evaluation using a commercial solver and a specialized classical chip for QUBO solving demonstrates the strengths of the two models. 
Our results show that both approaches yield fast and high-quality solutions 
and
improve over existing WFLO solutions. In particular, the QUBO model delivers high quality solutions very quickly while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
\end{abstract}
\begin{keyword}
	Wind Farm Layout Optimization (WFLO) \sep Quadratic Programming (QP) \sep Quadratic Unconstrained Binary Optimization (QUBO) \sep Digital Annealer
\end{keyword}

@@

\paragraph{Exact Optimization Methods} 
The WFLO problem can be solved  
to optimality, using
techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming (CP))~\cite{Zhang2014,turner2014new,donovan2005wind}.
These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
the best solution one can obtain for the given problem. Donovan was the first to solve WFLO using a
MILP approach based on the LS energy expression~\cite{donovan2005wind}. In~\cite{turner2014new}, the authors
approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well
compared to existing approximate and earlier exact solutions.

>>>>

\title{\mytitle}

%\subtitle{Extended Abstract}
%\author{}
%\institute{}
%\author{\myauthor}
%\institute{Department of Mechanical and Industrial Engineering, University of Toronto\\
%\email{sariks@mie.utoronto.ca, ecohen@mie.utoronto.ca, jcb@mie.utoronto.ca}}

@@

\begin{abstract} 
%Wind farms play an increasingly important
%role as a source of renewable energy and 
  A key aspect of the design of a wind farm is the wind farm layout optimization (WFLO) problem: given a wind farm site, decide the location of individual wind turbines to maximize energy production subject to proximity restrictions and wake-based interference among turbines. Given the pairwise wake interactions, it is natural
%While a number of heuristic and exact techniques have been proposed to solve WFLO, it is common
to model the energy objective as a quadratic function as, indeed, has been done in some existing optimization models. Recent advances in both optimization hardware and software have targeted quadratic constraints: commercial mixed integer linear solvers have been extended to address some quadratic problems and nascent specialized hardware, including quantum computing systems, have focused on solving Quadratic Unconstrained Binary Optimization (QUBO) problems. 
%It is a key problem 
%when designing a new wind farm, as it has a direct influence
%on the amount of the total produced energy. 
%In recent years there have been advances in both software technologies (e.g., the extension of commercial mixed integer linear programming solvers to handle quadratic problems) and nascent hardware development (e.g., quantum computing chips and specialized classical chips) that have relevance for the WFLO problem. In both cases, the advances have 
%Recent years saw 
%advances in software and hardware optimization technologies, such as Gurobi's quadratic solver 
%and Fujitsu's digital annealer (DA), respectively. 
%These technologies create an opportunity to improve 
%over current wind farm optimization solutions
%both in terms of solution quality and solve time. 
%In order to exploit these technologies, 
%we must represent the WFLO problem as 
%a quadratic optimization model. Specifically, in order 
%to apply a software-based optimization
%solution (Gurobi), 
In this paper, we 
introduce two novel quadratic programming models for WFLO: a quadratic constrained optimization problem (QCOP) with binary decision variables and a QUBO.
%Subsequently, we provide the unconstrained counterpart of the QCOP, 
%namely the quadratic unconstrained binary optimization (QUBO), which  
%is the underlying representation for quantum-inspired 
%optimization hardware such as the DA. 
A thorough empirical 
evaluation using a commercial solver and specialized QUBO hardware
%demonstrates the strengths of the two models. 
%Our results
show that both approaches yield fast, high-quality solutions 
that
improve the state of the art. In particular, the QUBO model delivers high quality solutions very quickly while the QCOP model can be used to find better solutions and provide quality guarantees over a longer run-time.
\end{abstract}
\begin{keyword}
	Wind Farm Layout Optimization (WFLO) \sep Quadratic Programming (QP) \sep Quadratic Unconstrained Binary Optimization (QUBO) \sep Digital Annealer
\end{keyword}

@@

\paragraph{Exact Optimization Methods} 
The WFLO problem can be solved  
to optimality, using
techniques from operations research (e.g., mixed-integer linear programming (MILP), quadratic programming (QP), and constraint programming (CP))~\cite{Zhang2014,turner2014new,donovan2005wind}.
These methods are \emph{exact}, in the sense that they guarantee that the returned solution is indeed
the best solution one can obtain for the given problem. Donovan was the first to solve WFLO using a
MILP approach based on the LS energy expression~\cite{donovan2005wind}. In~\cite{turner2014new}, the authors
approximated the SS energy expression using a quadratic and a linear approximation. The two resulting methods performed well
compared to existing approximate and earlier exact solutions.
\todo{Add the NLM approach - explain how it becomes a different problem \cite{ulku2019new}}
